\textbf{Hardware accelerator topologies in heterogeneous systems: an evaluation in the security domain}

### 2019-12-08 reports/paper.tex
write more

\begin{abstract}
  Hardware accelerators (HWACCs) have driven performance gains for
  domain specific applications which leverage highly optimized computation
  cores. However, the traditional accelerator access topology as an add-on to
  general purpose processors through shared memory space/bus, reveals performance
  issues as the number of HWACCs grow and systems become
  communication-bound.
  On the other hand, efficient development and research of HWACCs
  that are closer to a general purpose processor's datapath is seen as a
  prohibitive task due to the required microarchitecture knowledge and lack of
  tools to explore the design space. This has in turn hindered the exploration of
  such architectures for gains in either performance or security spectrums.
  RISC-V projects made available in the last few years provide a framework that
  greatly eases microarchitecture research, enabling further experimentation
  into security and performance aspects for heterogeneous platforms.
  This work uses a customized RISC-V SoC to explore performance trade-offs
  between different HWACC topologies (in-datapath, memory-mapped) for embedded
  systems in the security domain by implementing an IP core for cryptographic
  primitive acceleration.
  % results also go in here

@@

To assemble an entire SoC for cycle-accurate simulation, UC Berkeley's Chipyard~\cite{Chipyard}
is chosen.

@@

The hybrid crypto accelerator is distinct from the tightly-coupled topology in
the way that it has a DMA port for direct access to the coherent memory
hierarchy. This allows the accelerator to directly request data to be loaded and
stored, enabling two different modes of operation.

@@

@@

>>>>

\begin{abstract}
  Hardware accelerators (HWACCs) have driven performance gains for
  domain specific applications which leverage highly optimized computation
  cores. However, the traditional accelerator access topology as an add-on to
  general purpose processors through shared memory space/bus, reveals performance
  issues as the number of HWACCs grow and systems become
  communication-bound.
  On the other hand, efficient development and research of HWACCs
  that are closer to a general purpose processor's datapath is seen as a
  prohibitive task due to the required microarchitecture knowledge and lack of
  tools to explore the design space. This has in turn hindered the exploration of
  such architectures for gains in either performance or security spectrums.
  RISC-V projects made available in the last few years provide a framework that
  greatly eases microarchitecture research, enabling further experimentation
  into security and performance aspects for heterogeneous platforms.
  This work uses a customized RISC-V SoC to explore performance trade-offs
  between different HWACC topologies (in-datapath, memory-mapped) for embedded
  systems in the security domain by implementing an IP core for cryptographic
  primitive acceleration. We also show a hybrid accelerator topology for
  acceleration and/or secure memory storage which offers a speedup of 6x compared to a
  pure software implementation.
  % results also go in here

@@

To assemble an entire SoC for cycle-accurate simulation, UC Berkeley's Chipyard~\cite{Chipyard}
is chosen. Specifically, technologies used include:

\subsubsection{RISC-V cores}

There is already a variety of readily usable RISC-V cores out in the wild. To
study different HWACC topologies, UC Berkeley's implementations are chosen; we
select Rocket (in-order 5 stage pipeline, riscv64) and Berkeley out-of-order
machine (BOOM).

\subsubsection{RoCC interface}

The RoCC interface is an interface developed for the Rocket RISC-V core that
allows easy integration of HWACCs, effectively enabling the construction of
tightly-coupled accelerators. This interface is also supported in UC Berkeley's
BOOM implementation. In this paper, the terms tightly-coupled accelerator and
RoCC accelerator are used interchangeably when talking about the specific
implementations related to evaluation and results.

The RISC-V ISA specifies 4 different opcodes, \texttt{custom0-3} which are used for passing data
into and back the RoCC interface; with this mechanism, the HWACC is receives the
contents of up to two CPU registers and optionally passes back a word to be stored in any of
the CPU's registers as a result of the operation.

@@

The hybrid crypto accelerator is distinct from the tightly-coupled topology in
the way that it has a DMA port for direct access to the coherent memory
hierarchy. This allows the accelerator to directly request data to be loaded and
stored, enabling two different modes of operation. Note that due to the
architecture of the RoCC interface, the accelerator will have access to the
coherent memory hierarchy at the L2 cache level.

@@

Figure~\ref{fig:scatter} shows the relationship between number of instructions
retired by processor and total number of cycles for the benchmark to finish in
the in-order (rocket) core. Ultimately, the number of instructions retired is
correlated to code size; it is clear that for the accelerator configurations
that allow encryption in the CBC mode to be implemented, there is little
overhead incurred by the extra operations (XORs), which are carried out by
the CPU.

@@

The tools provided by the RISC-V community members enable computer architecture
exploration in a very powerful way. This allowed for the exploration of
different HWACC topologies; we found that there can be performance gains when
using tightly-coupled accelerators and security gains.

In order to achieve maximum performance/security, however, it is necessary to
fully customize the microarchitecture and toolchain to fully orchestrate secure
computations in the CPU's datapath, which will increase development costs.
